---
title: "Week 6 Lecture: Quarto pt. 2!"
author: "Natalie Goeler-Slough"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    warning: false
---

### Loading my libraries:

```{r}
#| message: false

library(palmerpenguins) #loading palmer penguins package with 'penguins' dataset
library(tidyverse)
library(kableExtra)
```

#### Loading data:
Dataset already included in palmerpenguins package; has data on penguins collected from Palmer Station LTER.
```{r}
head(penguins) #looking at head of penguins data to remind myself of available variables
```


#### Data analysis:

Making preliminary plot with penguins data:
```{r}
#| warning: false

plot1 <- penguins |>
  ggplot(aes(x = species, y = bill_length_mm)) +
  geom_violin() +
  geom_jitter(aes(color = island)) +
  labs(title = "Penguin Bill Length across Species & Islands",
       x = "Species", y = "Bill Length (mm)",
       caption = "Source: Palmer Station LTER",
       color = "Island")+
  theme_bw()
plot1
```

Now, practicing changing the size & alignment of the figure (half the width & centered)

```{r}
#| out-width: "50%" #making output 50% of the width size
#| fig-align: "center" #putting plot into the center
#| warning: false
plot1
```
Now, adding labels:
```{r}
#| label: penguin #now this plot is labeled as "penguin" and can be called
#| fig-cap: "This is my penguin figure"

plot1
```
Using kableExtra to make a table:
```{r}
penguins |>
  group_by(species) |>
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) |>
  kbl() # make it a kable table
```

Adding style to the table:
```{r}
penguins |>
  group_by(species) |>
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) |>
  kbl()  |> # make it a kable table
  kable_classic() |> # add a theme, kable classic
  row_spec(2, bold = TRUE, color = "white", background = "turquoise") |> # highlight row 2 in turquoise
  kable_styling(full_width = FALSE) # don't make the table so wide (making it narrower)
```

In-class activity: create code chunks with a ggplot figure and kbl() table, label and caption both, and then make a prose paragraph that references both.
```{r}
#| label: fig-penguin-plot
#| fig-cap: "Bill length across 3 penguin species on 3 islands."
plot1 #already made plot above (& assigned to name 'plot1'), using again here
```
```{r}
#| label: tbl-penguin-averages
#| tbl-cap: "Average penguin bill length by species and island."

penguins |>
  group_by(species, island) |>
  summarise(mean_bill_length = mean(bill_depth_mm, na.rm = TRUE)) |>
  kbl()
```

As shown in @fig-penguin-plot, bill lengths vary across species and island. Mean bill lengths across species and islands is summarized in @tbl-penguin-averages.
