---
title: "Week_7_Lecture"
author: "Natalie Goeler-Slough"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    warning: false
---

## Loading libraries
```{r}
library(tidyverse)
library(here)
library(sf)
library(maptiles)
library(tidyterra)
library(leaflet)
library(tidygeocoder)
library(ggspatial)
```

## Loading data
```{r}
ChemData <- read_csv(here("Week_7", "Data", "chemicaldata_maunalua.csv"))

glimpse(ChemData)
```

## Data analysis

### Tile and satellite basemaps with {maptiles}:

Using maptiles to get tile basemaps of Wailupe, Oahu -- setting a bounding box:
```{r}
# Define a bounding box around Wailupe, Oahu
# xmin/xmax = westmost/eastmost longitude
# ymin/ymax = southernmost/northernmost latitude
wailupe_bbox <- st_as_sfc(
  st_bbox(c(xmin = -157.772, ymin = 21.267,
            xmax = -157.752, ymax = 21.281),
          crs = 4326) # crs = 4326 means WGS84 — the lat/lon system your GPS uses
)
```

Downloading the tiles:
```{r}
#downloading the tiles
tiles_satellite <- get_tiles(
  wailupe_bbox,               # the area to download
  provider = "Esri.WorldImagery", # which tile source to use
  zoom = 16,                  # level of detail (3 = continent, 20 = building)
  crop = TRUE,                # clip to our exact bbox (vs. slightly larger)
  cachedir = here("Week_7", "Data", "tiles_cache") # save tiles to disk (caching, so it runs faster)
)
class(tiles_satellite)
```

Plotting with geom_spatraster_rgb(), a tidyterra function
```{r}
# geom_spatraster_rgb() is from {tidyterra}
# It reads the 3 color layers and renders them as a photo
ggplot() +
  geom_spatraster_rgb(data = tiles_satellite) +
  coord_sf(crs = 4326) +
  theme_minimal()
```
Deriving the bounding box (Bbox) from your Data, instead of hard-coding coordinates:
```{r}
# Step 1: convert your data frame to an sf object
# coords tells sf which columns hold longitude and latitude
ChemData_sf <- st_as_sf(ChemData,
                         coords = c("Long", "Lat"),
                         crs = 4326)

# Step 2: get the bounding box and add a small buffer
# the + c(...) pads all four sides by ~0.4 km so points aren't at the edge
bbox_chem <- st_bbox(ChemData_sf) + c(-0.004, -0.004, 0.004, 0.004)

# Step 3: download tiles for that extent
Map_satellite <- get_tiles(
  st_as_sfc(bbox_chem),
  provider = "Esri.WorldImagery",
  zoom = 15,
  crop = TRUE,
  cachedir = here("Week_7", "Data", "tiles_cache")
)
```

### Overlaying data + annotations with {ggspatial}:

Overlaying point data onto tile basemap:
```{r}
#plotting points from chemdata, colored by salinity (to see where freshwater vs seawater is)
ggplot() +
  geom_spatraster_rgb(data = Map_satellite) +
  geom_point(data = ChemData,
             aes(x = Long, y = Lat,
                 color = Salinity),
             size = 4) +
  scale_color_viridis_c() +
  coord_sf(crs = 4326) +
  theme_minimal()
```
Adding a scale bar and north arrow (from ggspatial package):
```{r}
ggplot() +
  geom_spatraster_rgb(data = Map_satellite) +
  geom_point(data = ChemData,
             aes(x = Long, y = Lat,
                 color = Salinity),
             size = 4) +
  scale_color_viridis_c() +
  annotation_scale(           # adding scale bar
    location = "bl",          # location: bl = bottom-left
    bar_cols = c("yellow", "white")) +
  annotation_north_arrow(     # adding north arrow
    location = "tl") +        # tl = top-left
  coord_sf(crs = 4326) +
  theme_minimal()
```
Polishing and saving:
```{r}
ggplot() +
  geom_spatraster_rgb(data = Map_satellite) +
  geom_point(data = ChemData,
             aes(x = Long, y = Lat, color = Salinity),
             size = 4) +
  scale_color_viridis_c(name = "Salinity (ppt)") +
  annotation_scale(location = "bl",
                   bar_cols = c("yellow", "white")) +
  annotation_north_arrow(location = "tl") +
  coord_sf(crs = 4326) +
  labs(title = "Maunalua Bay — Water Chemistry",
       caption = "Basemap: ESRI World Imagery") +
  theme_minimal()

ggsave(here("Week_7", "Output", "maunalua_map.pdf"), width = 8, height = 6)
```

### Interactive maps with {leaflet}:

Building my first leaflet map! Uses piping with |>
```{r}
leaflet() |>
  addProviderTiles(providers$Esri.WorldImagery) |>
  setView(lng = -157.762, lat = 21.274, zoom = 14) #sets initial view but you can zoom in/out and move
```

Adding a color palette to be able to plot with leaflet
```{r}
pal <- colorNumeric(
  palette = "viridis",            # any RColorBrewer or viridis palette name
  domain = ChemData$Salinity      # the full range of values to map
)

# Use it by calling pal(value) — it returns a hex color
pal(30)  # what color does salinity = 30 get?
pal(40)
```

Now adding data points with color:
```{r}
pal <- colorNumeric("viridis", domain = ChemData$Salinity)

leaflet(ChemData) |>           # pass data frame here
  addProviderTiles(providers$Esri.WorldImagery) |>
  addCircleMarkers(
    lng = ~Long, lat = ~Lat,   # which columns are coordinates (longitude and latitude, with ~ indicating a column)
    color = ~pal(Salinity),    # apply our color function to Salinity, with the palette from before
    radius = 6,
    fillOpacity = 0.9,
    popup = ~paste("Salinity:", Salinity, "ppt",  # click a point to see this (to see salinity)
                   "<br>Site:", Site)) |>
  addLegend(
    pal = pal,                 # the same color function
    values = ~Salinity,        # which column drives the legend
    title = "Salinity (ppt)",
    position = "bottomright")
```


### Geocoding with {tidygeocoder}:
(Geocoding = converting a place name or address into latitude and longitude coordinates)

```{r}
geo("University of Hawaii at Manoa", method = "osm") #"osm" = open street maps, alternate: "census" = best for US street addresses

geo("Pearl Harbor, Oahu, Hawaii", method = "osm") #doesn't come up, likely will have to be more specific

```
Geocoding a whole dataframe:
```{r}
oahu_sites <- tribble(
  ~place,
  "University of Hawaii at Manoa",
  "Honolulu International Airport",
  "Pearl Harbor National Memorial, Hawaii",
  "Diamond Head State Monument, Hawaii"
)

oahu_geocoded <- oahu_sites |>
  geocode(place, method = "osm") # geocode the 'place' column

oahu_geocoded
```
Mapping geocoded locations using leaflet:
```{r}
leaflet(oahu_geocoded) |>
  addProviderTiles(providers$Esri.WorldImagery) |>
  addMarkers(
    lng = ~long, lat = ~lat,   # tidygeocoder always returns 'lat' and 'long'
    label = ~place)            # hover to see place name
```

### Today's totally awesome package: {emojifont}!
```{r}
library(emojifont)

search_emoji("smile")

ggplot() +
  geom_emoji("smiley_cat",
             x = 1:5, y = 1:5,
             size = 10)
```

