---
title: "Week_7_Class"
author: "Natalie Goeler-Slough"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    warning: false
---

## Loading libraries
```{r}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
```


## Loading data
```{r}
# Population in California by county
popdata <- read_csv(here("Week_7", "Data", "CApopdata.csv"))
head(popdata)

# Seastar counts at field sites in California
stars <- read_csv(here("Week_7", "Data", "stars.csv"))
head(stars)
```


## Data analysis: making maps!

### Pulling polygon boundaries:
```{r}

world <- map_data("world") 
head(world)

usa <- map_data("usa")
head(usa)

states <- map_data("state")
head(states)

italy <- map_data("italy")
head(italy)

counties <- map_data("county")
head(counties)
```

### Making a map of the world!
```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group))
```
If you forget group=group:
```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat))
```
Adding color, lines, & changing theme:
```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region), #filling each country/region
               color = "black") + # outlining each country/region
  guides(fill="none") + #gets rid of fill legend
  theme_minimal()
```

Making the ocean blue:
```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region), #filling each country/region
               color = "black") + # outlining each country/region
  guides(fill="none") + #gets rid of fill legend
  theme(panel.background = element_rect(fill="lightblue"))
```

Changing map projection to mercator:
```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region),
               color = "black") +
  theme_minimal() +
  guides(fill = "none") +
  theme(panel.background =
          element_rect(fill = "lightblue")) +
  coord_map(projection = "mercator", #mercator projection
            xlim = c(-180, 180))
```

Changing map projection to sinusoidal:
```{r}
ggplot() +
  geom_polygon(data = world,
               aes(x = long, y = lat,
                   group = group,
                   fill = region),
               color = "black") +
  theme_minimal() +
  guides(fill = "none") +
  theme(panel.background =
          element_rect(fill = "lightblue")) +
  coord_map(projection = "sinusoidal",
            xlim = c(-180, 180))
```

### Making a map of just California:
(and mapping each county)
```{r}
#filtering states dataset for only CA
CA_data <- states |>
  filter(region == "california")
head(CA_data)

CA_counties <- counties |>
  filter(region == "california")

ggplot()+
  geom_polygon(data = CA_data,
               aes(x=long, y=lat,
                   group = group),
               color = "purple", 
               fill = "magenta")+
  geom_polygon(data=CA_counties,
               aes(x=long, y=lat,
                   group = group,
                   fill=subregion),
               color = "black")+
  guides(fill = "none") +
  coord_map()
```
### Adding multiple layers of data:
#### Plotting the population of each county in CA
```{r}
head(counties)[1:3, ]
head(popdata)
```
Wrangling data:
```{r}
CApop_county <- popdata |>
  select("subregion" = County, Population) |>   # rename to match counties
  inner_join(counties) |>
  filter(region == "california")
head(CApop_county)
```
Making map of CA population by county:
```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  coord_map() +
  theme_void()
```
Log scale for easier interpretation:
```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  coord_map() +
  theme_void() +
  scale_fill_gradient(transform = "log10")
```
Adding a layer, overlaying sea star sites:
```{r}
head(stars)

ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  geom_point(data = stars,
             aes(x = long,
                 y = lat)) +
  coord_map() +
  theme_void() +
  scale_fill_gradient(transform = "log10")
```

Now changing size of points to reflect sea star count, fixing legend, and changing point color:
```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long, y = lat,
                   group = group,
                   fill = Population),
               color = "black") +
  geom_point(data = stars,
             aes(x = long, y = lat,
                 size = star_no),
             color = "red") + #sizing points by sea star count
  coord_map() +
  theme_void() +
  scale_fill_gradient(transform = "log10") +
  labs(size = "# stars/m²") #fixing legend label

ggsave(here("Week_7", "Output", "CApopstars.pdf"))
```

### Now, using {sf} package:
```{r}
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r}
world_sf <- ne_countries(scale = "medium", #ne = natural earth, scale = scale of map
                         returnclass = "sf")
ggplot(world_sf) +
  geom_sf(fill = "lightblue",
          color = "white") +
  theme_void()
#don't need group, or lat/long columns! {sf} handles it all
```

Projections with coord_sf():
```{r}
ggplot(world_sf) +
  geom_sf(fill = "lightgray", color = "white") +
  coord_sf(crs = "+proj=robin") + # Robinson projection
  theme_void() +
  labs(title = "Robinson Projection") #adding title for projection selection
```

Totally awesome R package: {ggdogs}!
```{r}
library(ggdogs)
ggplot(mtcars) +
  geom_dog(aes(mpg, wt), dog = "pug", size = 5)
```

