---
title: "Week 7 Homework: Mapping"
author: "Natalie Goeler-Slough"
format:
  html:
    code-fold: true
    code-summary: "Show code"
    warning: false
    theme: lux
---

## Loading libraries
```{r}
library(readr)
library(leaflet)
library(tidyverse)
library(here)
library(sf)
library(maptiles)
library(tidyterra)
library(ggspatial)
library(leaflegend)
```


## Loading data
Using dataset on meteorites! Where they fell & when they fell. Comes from the the Meteoritical Society by way of NASA.
```{r}
meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2019/2019-06-11/meteorites.csv")

glimpse(meteorites)
```

## Data analysis:

### Making map of meteorites!

First, I'm defining reference mass values and setting up calculated values for my legend.
```{r}
#first, need to define reference mass values and labels for plotting the legend
legendmasses <- c(1000, 10000, 100000, 1000000, 10000000) #reference mass values (in grams, matching meteorite dataframe)
masslabels <- c("1 kg", "10 kg", "100 kg", "1,000 kg", "10,000 kg") #labels for reference mass values
diameters <- 2 * (sqrt(legendmasses) * 0.005) #calculating diameters for legend based on scaling factor I'm using to plot
named_breaks <- setNames(diameters, masslabels) #making a vector with calculated diameter values and labels of corresponding masses
```

Now, I'm plotting! Mapping meteorite sightings across the world, sizing points based on the mass of the meteorite. Also making the map interactive such that you can click on a point to see the meteor name and year that the meteor was found.

### Global Meteorite Mass Distribution:
```{r}
#| label: fig-meteormap
#| fig-cap: "Global Meteorite Mass Distribution. *Click on individual points to view meteorite names and discovery years.* Source: NASA."
leaflet(meteorites) |>
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012) |> #using a fun earth at night map base :) 
  setView(lng = 45.0, lat = 30.0, zoom = 2.5) |> #setting initial zoom to be around Europe/Asia/Africa
  addCircleMarkers(
    lng = ~long, lat = ~lat,
    radius = ~sqrt(mass) * 0.005, #scaling circles to be the mass of the meteor (sqrt and making smaller to scale onto map)
    color = "red",
    fillOpacity = 0.8,
    popup = ~paste("Name:", name,  # click a point to see the name of the meteor
                   "<br>Year:", year)) |>  #also displays the year the meteor was found
  addLegendSize( #adding legend for sizes
    values = diameters, #using calculated diameters from above, based on scaling factor
    breaks = named_breaks, #breaks by selected mass values above, labeled with actual weight in kg
    baseSize = mean(diameters),
    position = "bottomleft",
    title = "Meteorite Mass",
    shape = "circle",
    color = "red",
    fillColor = "red",
    fillOpacity = 0.8
  ) 
```



